Thermodynamic fragments for magic states in quantum computation

### 2021-02-25 magic_paper.tex
Start of stabilizer section & first appendices, I am unsure of how structure unital section

Let $\{\ket{k}\}_{k \in \mathbb{Z}_d}$ be the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$, with $d$ an odd prime. 
Since the field has character $d$, addition and multiplication on the field are always considered modulo $d$.
The Hilbert space of any system associated with this scheme is $\hd \coloneqq \spanv{\ket{k}: k \in \mathbb{Z}_d}$.

The generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}, \label{eq:zpauli}
\end{align}
where $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is modulo $d$. The Hilbert space $\hd$ is associated with a phase space $\cal{P}_d = \mathbb{Z}_d \times \mathbb{Z}_d$, where every point $\bmx \coloneqq (x, p)$ corresponds to a displacement operator, defined as
\begin{equation}\label{eq:ddef}
    D_{\bmx} \coloneqq \tau^{x p} X^{x} Z^{p},\ (x, p) \in \cal{P}_d.
\end{equation}
The phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity.
For a composite system with product dimension $d = d_1 \dots d_n$ and Hilbert space, $\hd = \hd[d_1] \otimes \dots \otimes \hd[d_n]$, the displacement operators are defined as
\begin{equation}\label{eq:composited}
    D_{\bmx} \coloneqq D_{(x_1, p_1)} \otimes \dots \otimes D_{(x_n, p_n)},
\end{equation}
where 
\begin{align*}
	\bmx \coloneqq (x_1, p_1, x_2, p_2, \dots, x_n, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n} \eqqcolon  \cal{P}_d,
\end{align*}
denotes the phase space point for an $n$--copy system.

@@

The phase-point operators form an orthogonal Hermitian operator basis as shown in~\cref{app:wigner}.
Therefore, any quantum state $\rho \in \cal{B}(\hd)$ can be expressed as a linear combination of them,
\begin{equation}
    \rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz},
\end{equation}
where the coefficient vector $\W{\rho}$ is the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmx]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmx} \rho].
\end{equation}

@@

To formally state majorization results, we first denote by $\stoch$ the set of $(d \times d)$ stochastic matrices that preserve the probability vector $\bmd$.
Specifically, for any $S \in \stoch$, all matrix elements are non-negative, all rows sum to $1$ and $S\bmd = \bmd$.
\iffalse
\begin{enumerate}
    \item $S_{ij} \geq 0$ for all $i, j \in \mathbb{Z}_d$;
    \item $\sum\limits_{j=1}^n S_{ij} = 1$ for all $i \in \mathbb{Z}_d$;
    \item $S\bmd = \bmd$.
\end{enumerate}
\fi
The set $\stoch$ forms a group under matrix multiplication for all $\bmd$ with positive components.

Majorization exactly describes quantum thermodynamics in the absence of coherence.
The use of majorization in this setting provides vital intuition for our purposes.
At any given temperature $\beta^{-1}$, the thermal state $\gamma_\beta$ is thermodynamically the most disordered state. 
Thermal operations are defined as operations that cannot extract energy from the Gibbs state, $\E(\gamma_\beta) = \gamma_\beta$.
Convertibility between states via thermal operations is equivalent to a stochasticity condition on the energy level populations of the states \nick{CITE}.
Informally, there exists a thermal operation $\E$ such that $\tau = \E(\rho)$ if and only if there exists a matrix $S \in \stoch$ such that $\bm{q} = S\bm{p}$, where $\bm{q}, \bm{p}$ and $\bmd$ and the energy level population vectors of $\tau, \rho, \gamma_\beta$ respectively.

@@

    Consider the piecewise linear curve obtained by joining the points $\{(0,0)\} \cup \{ (x_k, \lc{\bmz}{\bmd}(k)) \}_{k=1,\dots,d}$, where
    \begin{equation}\label{eq:lorenz}
        (x_k, \lc{\bmw}{\bmd}(k)) \coloneqq \left( \frac{1}{D}\sum_{i=1}^k d_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right).
    \end{equation}
    We define the set of points on this curve, $\lc{\bmw}{\bmd}(x),\ x \in [0,1]$, as the \emph{Lorenz curve} of vector $\bmw$ with respect to $\bmd$.
\end{definition}
Components $x_k$ are rescaled by $D$ so that comparison of curves with unequal dimensions is possible.
In fact, the Lorenz curves $\lc{\bmw}{\bmd}$ and $\lc{\bmw \otimes \bmd}{\bmd \otimes \bmd}$, where $\otimes$ denotes the Kronecker product, are identical.
Furthermore, a Lorenz curve $\lc{\bmw}{\bmd}(x)$ is always concave in $x$, since it consists of $d$ line segments each with slope $(\widetilde{\bmw}^\downarrow)_i$ for $i=1,\dots,d$ which is non-increasing by definition.
Finally, the points on the interior of the Lorenz curve connecting line segments of different slopes are called \emph{elbows}.

@@

We focus on the distillation process
\begin{equation}\label{eq:sudist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \gamma_\beta^{\otimes (n-n')},
\end{equation}
where the noisy Strange state is given in~\cref{eq:noisysn}, but more general states can be handled in exactly the same manner.
Notice that tensoring in copies of $\gamma_\beta$ does not affect the process, since the state is free.
In the following result, all copies $n, n', n - n'$ are even with $n > n'$, and $\epsilon'=0$, but the bounds are easily generalised to odd numbers of copies and $\epsilon'$ such that $0 < \epsilon' < \epsilon$. 
Finally, we assume a \nick{generous} range of initial noise levels, $\epsilon \leq 3/7$, so that the largest Wigner component of the 1-copy noisy magic state is negative.

>>>>

Let $\{\ket{k}\}_{k \in \mathbb{Z}_d}$ be the standard computational basis, defined over $\mathbb{Z}_d = \{ 0, 1, \dots,d-1 \}$, with $d$ an odd prime. 
Since the field has character $d$, addition and multiplication on the field are always considered modulo $d$.
The Hilbert space of any system associated with this scheme is $\cal{H}_d \coloneqq \spanv{\ket{k}: k \in \mathbb{Z}_d}$.

The generalised Pauli matrices $X, Z$ can be defined by their respective roles as shift and phase operators,
\begin{align}
    X \ket{k} &= \ket{k + 1} \label{eq:xpauli}\\
	Z \ket{k} &= \omega^k \ket{k}, \label{eq:zpauli}
\end{align}
where $\omega \coloneqq e^{2\pi i/d}$ is the $d$-th root of unity and addition is modulo $d$. The Hilbert space $\cal{H}_d$ is associated with a phase space $\cal{P}_d = \mathbb{Z}_d \times \mathbb{Z}_d$, where every point $\bmx \coloneqq (x, p)$ corresponds to a displacement operator, defined as
\begin{equation}\label{eq:ddef}
    D_{\bmx} \coloneqq \tau^{x p} X^{x} Z^{p},\ (x, p) \in \cal{P}_d.
\end{equation}
The phase factor $\tau \coloneqq -\omega^{1/2}$ ensures unitarity.
For a composite system with product dimension $d = d_1 \dots d_n$ and Hilbert space, $\cal{H}_d = \cal{H}_{d_1}] \otimes \dots \otimes \cal{H}_{d_n}]$, the displacement operators are defined as
\begin{equation}\label{eq:composited}
    D_{\bmx} \coloneqq D_{(x_1, p_1)} \otimes \dots \otimes D_{(x_n, p_n)},
\end{equation}
where 
\begin{align*}
	\bmx \coloneqq (x_1, p_1, x_2, p_2, \dots, x_n, p_n) \in \cal{P}_{d_1} \times \dots \times \cal{P}_{d_n} \eqqcolon  \cal{P}_d,
\end{align*}
denotes the phase space point for an $n$--copy system.

@@

The phase-point operators form an orthogonal Hermitian operator basis as shown in~\cref{app:wigner}.
Therefore, any quantum state $\rho \in \cal{B}(\cal{H}_d)$ can be expressed as a linear combination of them,
\begin{equation}
    \rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz},
\end{equation}
where the coefficient vector $\W{\rho}$ is the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmx]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmx} \rho].
\end{equation}

@@

To formally state majorization results, we first denote by $\stoch$ the set of $(d \times d)$ stochastic matrices that preserve the probability vector $\bmd$.
Specifically, for any $S \in \stoch$, all matrix elements are non-negative, all rows sum to $1$ and $S\bmd = \bmd$.
The set $\stoch$ forms a group under matrix multiplication for all $\bmd$ with positive components.

Majorization describes quantum thermodynamics exactly in the absence of quantum coherence.
The use of majorization in this setting provides vital intuition for our purposes.
At any given temperature $\beta^{-1}$, the thermal state $\gamma_\beta$ is thermodynamically the most disordered state. 
Thermal operations are defined as operations that cannot extract energy from the Gibbs state, $\E(\gamma_\beta) = \gamma_\beta$.
Convertibility between states via thermal operations is equivalent to a stochasticity condition on the energy level populations of the states \nick{CITE}.
Informally, there exists a thermal operation $\E$ such that $\tau = \E(\rho)$ if and only if there exists a matrix $S \in \stoch$ such that $\bm{q} = S\bm{p}$, where $\bm{q}, \bm{p}$ and $\bmd$ and the energy level population vectors of $\tau, \rho, \gamma_\beta$ respectively.
Formal statements of the above can be found \nick{in the appendix}.

@@

    Consider the piecewise linear curve obtained by joining the points $\{(0,0)\} \cup \{ (x_k, \lc{\bmw}{\bmd}(k)) \}_{k=1,\dots,d}$, where
    \begin{equation}\label{eq:lorenz}
        (x_k, \lc{\bmw}{\bmd}(k)) \coloneqq \left( \frac{1}{D}\sum_{i=1}^k d_{\pi(i)}, \sum_{i=1}^k w_{\pi(i)} \right).
    \end{equation}
    We define the set of points on this curve, $\lc{\bmw}{\bmd}(x),\ x \in [0,1]$, as the \emph{Lorenz curve} of vector $\bmw$ with respect to $\bmd$.
\end{definition}
Components $x_k$ are rescaled by $D$ so that comparison of curves with unequal dimensions is possible.
In fact, the Lorenz curves $\lc{\bmw}{\bmd}$ and $\lc{\bmw \otimes \bmd}{\bmd \otimes \bmd}$, where $\otimes$ denotes the Kronecker product, are identical.
Furthermore, a Lorenz curve $\lc{\bmw}{\bmd}(x)$ is always concave in $x$, since it consists of $d$ line segments each with slope $(\widetilde{\bmw}^\downarrow)_i$ for $i=1,\dots,d$ which is non-increasing by definition.
Finally, the points on the interior of the Lorenz curve connecting line segments of different slopes are called \emph{elbows}.

@@

We focus on the distillation process
\begin{equation}\label{eq:stdist}
	\rho_{\rm{S}}(\epsilon)^{\otimes n} \longrightarrow \rho_{\rm{S}}(\epsilon')^{\otimes n'} \otimes \gamma_\beta^{\otimes (n-n')},
\end{equation}
where the noisy Strange state is given in~\cref{eq:noisysn}, but more general states can be handled in exactly the same manner.
Notice that tensoring in copies of $\gamma_\beta$ does not affect the process, since the state is free.
In the following result, all copies $n, n', n - n'$ are even with $n > n'$, and $\epsilon'=0$, but the bounds are easily generalised to odd numbers of copies and $\epsilon'$ such that $0 < \epsilon' < \epsilon$. 
Finally, we assume a \nick{generous} range of initial noise levels, $\epsilon \leq 3/7$, so that the largest Wigner component of the 1-copy noisy magic state is negative.
